Modelling the emergence of whisker barrels

### 2020-07-02 paper/briefpaper/brief.tex
Updates to a combined Fig 4

% This figure is text width
\begin{center}
  \includegraphics[width=\linewidth]{./Fig4.png}
  \captionof{figure}{Guidance
    fields and emergent barrel pattern in a Fgf8 misexpression experiment
    (c.f. \citealp{assimacopoulos_fibroblast_2012}), simulated by reflecting
    $\rho_1$ from Fig.\,\ref{fig:main}A at the join. \cmnt{The model
      parameters match those in Fig.\,\ref{fig:main}C. Scale
      bars are 1\,mm.}}
  \label{fig:fgf8}
\end{center}

\MPthreePred{To investigate the response of the model to environmental
  manipulation, we simulated a whisker deprivation experiment. In the first
  few days after birth, the upper region of the rat's cortex develops into
  distinct layers I--IV (layers V and VI are already distinct at
  birth). During this critical period, removal of the mystacial whiskers by
  electrocauterization, plucking or trimming leads to observable changes in
  brain structures, including the barrel field in cortical layer
  IV~\citep{jeanmonod_mouse_1981}. Amoung other changes, barrels associated
  with deprived whiskers have a smaller area. This is found whether the animal
  is deprived of an individual whisker~\citep{kossut_effects_1992}, or an
  entire row~\citep{loos_somatosensory_1973}, and demonstrates that
  thalamocortical activity contributes to the developmental process. We
  simulated the trimming of the individual whisker C3 during this critical
  period by reducing the amount by which its thalamocortical projections
  compete for branching. This was achieved by reducing the value of $\epsilon$
  in Eq.\,\ref{eq:comp} for the C3 projection, leaving other simulation
  parameters unchanged. The reduced competition for branching limited the area
  over which the C3 projection could make connections and its resulting barrel
  size was diminished (Fig.\,\ref{fig:whiskertrim}A). Its neighbours, whose
  ability to compete for branching was unchanged, were able to increase in
  size compared to the control simulation presented in Fig\,\ref{fig:main}C
  (Fig.\,\ref{fig:whiskertrim}B). The C3 barrel disappeared altogether if its
  competition parameter was reduced below a half of the original value.}

\begin{center}
  \includegraphics[width=\linewidth]{./Fig5.png}
  \captionof{figure}{\MPthreePred{Simulating whisker trimming by a reduction
      of the competition parameter, $\epsilon$, in the expression $\chi_i$
      (Eq.\,\ref{eq:comp}) for the C3 thalamocortical projection. For C3,
      $\epsilon$ is replaced by $m \epsilon$, where $m$ is a scalar
      multiplier. For the rest of the projections, $\chi_i$ is left unchanged
      and all other parameters are identical to those reported in
      Fig.\,\ref{fig:main}C. \textbf{A} The pattern formed after 30000 steps
      for $m=0.64$. Barrel C3 is reduced in size as its axon branching
      competes less strongly with its neighbours, whose area is slightly
      increased. \textbf{B} Individual barrel area plotted vs.~$m$
      shows that as $m$ is reduced, the area of C3 (blue squares) goes down,
      disappearing for $m\approx0.5$, whereas the average area of neighbouring
      barrels (triangles) increases.}}
  \label{fig:whiskertrim}
\end{center}

>>>>

% This figure is text width
\begin{center}
  \includegraphics[width=\linewidth]{./Fig4.png}
  \captionof{figure}{
    \MPthreePred{
      Two simulated whisker barrel manipulations.  Guidance fields
      (\textbf{A}) and emergent barrel pattern (\textbf{B}) in a Fgf8
      misexpression experiment
      (c.f.~\citealp{assimacopoulos_fibroblast_2012}), simulated by reflecting
      $\rho_1$ from Fig.\,\ref{fig:main}A at the join of the original boundary
      with its mirror. The model parameters match those in
      Fig.\,\ref{fig:main}C. \textbf{C} Simulating whisker trimming by a
      reduction of the competition parameter, $\epsilon$, in the expression
      $\chi_i$ (Eq.\,\ref{eq:comp}) for the C3 thalamocortical projection. For
      C3, $\epsilon$ is replaced by $m\,\epsilon$, where $m$ is a scalar
      multiplier. For the rest of the projections, $\chi_i$ is left unchanged
      and all other parameters are identical to those reported in
      Fig.\,\ref{fig:main}C. The pattern formed after 30000 steps with
      $m=0.64$. Barrel C3 is reduced in size as its axon branching competes
      less strongly with its neighbours, whose area is slightly
      increased. \textbf{D} Individual barrel area plotted vs.~$m$ shows that
      as $m$ is reduced, the area of C3 (blue squares) goes down, disappearing
      for $m\approx0.5$, whereas the average area of neighbouring barrels
      (triangles) increases. Scale bars are 1\,mm.
    }
  }
  \label{fig:fgf8}
\end{center}

\MPthreePred{To investigate the response of the model to environmental
  manipulation, we simulated a whisker deprivation experiment. In the first
  few days after birth, the upper region of the rat's cortex develops into
  distinct layers I--IV (layers V and VI are already distinct at
  birth). During this critical period, removal of the mystacial whiskers by
  electrocauterization, plucking or trimming leads to observable changes in
  brain structures, including the barrel field in cortical layer
  IV~\citep{jeanmonod_mouse_1981}. Amoung other changes, barrels associated
  with deprived whiskers have a smaller area. This is found whether the animal
  is deprived of an individual whisker~\citep{kossut_effects_1992}, or an
  entire row~\citep{loos_somatosensory_1973}, and demonstrates that
  thalamocortical activity contributes to the developmental process. We
  simulated the trimming of the individual whisker C3 during this critical
  period by reducing the amount by which its thalamocortical projections
  compete for branching. This was achieved by reducing the value of $\epsilon$
  in Eq.\,\ref{eq:comp} for the C3 projection, leaving other simulation
  parameters unchanged. The reduced competition for branching limited the area
  over which the C3 projection could make connections and its resulting barrel
  size was diminished (Fig.\,\ref{fig:fgf8}C). Its neighbours, whose
  ability to compete for branching was unchanged, were able to increase in
  size compared to the control simulation presented in Fig\,\ref{fig:main}C
  (Fig.\,\ref{fig:fgf8}D). The C3 barrel disappeared altogether if its
  competition parameter was reduced below a half of the original value.}
